---
title: "Figures for the paper ..."
author: Detlef Groth
date: 2023-02-05
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hermanussen et. al. figures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<center>[hgraph module](hgraph.html) - [paper figures](figures.html)</center>

## Introduction

Below you find the code for the paper ...

```{r}
x=1
print(x)
```

Season pairings:

```{r}
library(hanna)
simul_pairings(LETTERS[1:4])
```

Play a single round:

```{r}
res=simul_season(LETTERS[1:6])
res
```

Let's now conver this to an undirected graph where edges are drawn between
teams which draw against each other:

```{r}
U = simul_graph(res$M,mode='draw')
U
```

Now a directed graph where an edge is drawn from the winning team to the loosing team in a match.

```{r fig.width=8,fig.height=4,out.width=1000}
D = simul_graph(res$M,mode='win')
D
par(mfrow=c(1,2),mai=rep(0.2,4))
hgraph$plot(D,vertex.color=hgraph$colors(D))
hgraph$plot(U)
```

## Model comparison

```{r fig.width=14,fig.height=18,out.width=900,eval=FALSE}
pdf("~/Downloads/2023-02/networks.pdf",width=14,height=14)
set.seed(123)
par(mfrow=c(4,4),mai=c(0.0,0.5,0.5,0.0))
colors=c('grey80','#ff9999','#bb5555')
n=12
for (mode in c('null','memory5','gain','keystone')) {
    for (i in 1:30) {
        if (i == 1) {
            if (mode == "keystone") {
                res=simul_season(LETTERS[1:n],model=mode,token=c(24,24,rep(12,n-2)))
            } else if (mode=="memory1") {
               res=simul_season(LETTERS[1:n],model=mode,memory.length=1)
            } else if (mode=="memory5") {
               res=simul_season(LETTERS[1:n],model=mode,memory.length=5)
            } else {
               res=simul_season(LETTERS[1:n],model=mode)
            }
        } else {
            if (mode=="memory1") {
               res=simul_season(LETTERS[1:n],model=mode,memory.length=1,memory=res$memory)
            } else if (mode=="memory5") {
               res=simul_season(LETTERS[1:n],model=mode,memory.length=5,memory=res$memory)
            }  else {
                res=simul_season(LETTERS[1:12],token=res$token,model=mode)
            }
        }  
        D = simul_graph(res$M,mode='win') 
        cols=simul_colors(D,col=colors)
        if (i %in% c(1,5,10,30)) {
            simul_plot(D,vertex.color=cols,vertex.size=0.8)
            if (i == 1) {
                if (mode == "gain") {
                   mtext("winlose",side=2,cex=1.5,line=0)
                } else {
                    mtext(mode,side=2,cex=1.5,line=0)
                }   
            }
        
            if (mode == "null") {
                mtext(i,side=3,line=0,cex=1.5)
            }
        }
    }   
}
dev.off()
```

## Triad counting

```{r fig.width=14,fig.height=14,out.width=1000,eval=FALSE}
set.seed(128) 
pdf("~/Downloads/2023-02/triads.pdf",width=14,height=14)
par(mfcol=c(4,4),omi=c(0.5,0.7,0.5,0.3),mai=rep(0.2,4))  
for (i in c(1,5,10,30)) {
    res.df=simul_compare(n=10,seasons=i) 
    for (mod in c("null", "memory5","gain","keystone")) {  
       #rest=t(scale(t(res.df[res.df$model==mod,1:5]))) 
       rest=t(apply(res.df[res.df$model==mod,1:5],1,function(x) { (x/sum(x))*100} ))
       #rest=cbind(rest,res.df[,6])
       boxplot(rest,ylim=c(0,100),axes=FALSE) 
       lines(1:5,apply(rest,2,median)) 
       if (i == 1) {
          if (mod == "gain") {
            mtext("winlose",side=2,cex=1.5,line=3)
          } else {
            mtext(mod,side=2,cex=1.5,line=3)
          }
       }
       box()
       if (mod == "null") {
           mtext(i,side=3,cex=1.5,line=3)
       }
       if (i == 1) {
             axis(2,cex.axis=1.5)

       }
       if (mod == "keystone") {
         axis(1,labels=colnames(rest),at=1:5,cex.axis=1.5)
       }
        if (i == 1 & mod == "null") {
            res.all=cbind(res.df,iter=rep(i,nrow(res.df)))
        } else {
            res.all=rbind(res.all,cbind(res.df,iter=rep(i,nrow(res.df))))
        }
    } 
}
dev.off()
```

## Measure sd and mean for the average path lengths

```{r eval=FALSE}
agg1=with(res.all,aggregate(pls, by=list(model,iter), mean,na.rm=TRUE))
agg2=with(res.all,aggregate(pls, by=list(model,iter), sd,na.rm=TRUE))
agg3=with(res.all,aggregate(wls, by=list(model,iter), mean,na.rm=TRUE))
agg4=with(res.all,aggregate(wls, by=list(model,iter), sd,na.rm=TRUE))
head(agg1,n=20)
head(agg2,n=20)
agg=cbind(agg1,agg2$x,agg3$x,agg4$x)
colnames(agg)=c('model','seasons','mean.apl','sd.apl','mean.wapl','sd.wapl')
head(agg)
```


```{r eval=FALSE}
library(knitr)
knitr::kable(agg)
write.table(agg,file="~/Downloads/2023-02/path-lengths.tab",sep="\t",quote=FALSE)
```

## EOF



