---
title: "Short introduction into the hgraph module"
author: Detlef Groth, Detlef Groth, University of Potsdam, Germany
date: 2023-02-09
output: function () { rmarkdown::html_vignette(toc=TRUE,css="style.css") }
vignette: >
  %\VignetteIndexEntry{hgraph introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
include-before: |
    <center>[hgraph module](hgraph-tutorial.html) - [simul module](simul-tutorial.html) -
    [paper figures](figures.html)</center>
  
---

## Introduction

The `hgraph` object offers a few functions to deal with adjacency matrices
which are for instance created from the `simul` module. This module is a
simple R environment and can be inspected therefor using the `ls` function.
  

```{r}
library(hanna)  
ls(hgraph)
```

There are as well a few private functions which should not be used directly be the programmer.

```{r}
ls(hgraph,all=TRUE)
```

These private functions are starting with a dot at the beginning.

## Creating an example graph 

There is as well a function which can be used to create some example graphs.
Here an example were we create a so called Werner graph.

```{r fig.width=9,fig.height=6}
W=hgraph$graph()
cols=hgraph$colors(W)
hgraph$plot(W,vertex.color=cols)
```

The colors are indicating nodes with only outgoing edges in red, nodes with
in- and outgoing edges in blue and nodes with in- and outgoing edges in grey.

Let's create a larger random graph here:

```{r,fig.width=9,fig.height=6,out.width=800}
set.seed(124)
R=hgraph$graph(type="random",nodes=10, edges=14)
cols=hgraph$colors(R)
hgraph$plot(R,vertex.color=cols)
```

## Graph conversion

You can as well create undirected graphs, where the edges have no direction
out of directed graphs.

```{r fig.width=9,fig.height=6,out.width=800}
R
U=hgraph$d2u(R)
U
lay=hgraph$layout(U)
par(mfrow=c(1,2))
plot(R,vertex.color=cols,layout=lay)
plot(U,layout=lay)
```

## Graph information

We can extract various information about graphs and their nodes. For instance
how many edges has every node using the degree function or what is the length
of the shortest paths between all nodes. Here some example code:


```{r}
hgraph$degree(U)
hgraph$shortest_paths(U)
```

For an graph with several components the distance can be as well `Inf(inite)`.
Let's remove the edge between the nodes G and E for example and then
recalculate the distances.

```{r}
U2=U
U2[c('G','E'),c('G','E')]=0
hgraph$shortest_paths(U2)
```

If you need more functionality for working with graphs I suggest you use the
nice `igraph` package which is available for R.



