---
title: "Example simulations for the hanna package using the simul module"
author: Detlef Groth, University of Potsdam, Germany
date: 2023-02-13
vignette: >
  %\VignetteIndexEntry{Simul module tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<center>[hgraph module](hgraph-tutorial.html) - [simul module](simul-tutorial.html) -
[paper figures](figures.html)</center>

## Introduction

The hanna package contains two modules, environments for doing winner looser
simulations and evaluating the resulting network structure.

The simul environment provides the following methods:

* _simul$pairings_ - create round pairings for a single season
* _simul$season_ - create matches for everyone against everyone using the given model and evaluate the results
* _simul$graph_ - create a adjacency matrix out of the results for a match season
* _simul$compare_ - compare the different models for a certain number of seasons, does a lot of iterations to compare the models several times

## simul$pairings

This just allows to create matches where in every round the agents get a new partner, here an example for a small group of 6 agents giving all together 15 matches:

```{r}
library(hanna)
set.seed(123)
simul$pairings(LETTERS[1:6])
```

The function is just needed internally in the `simul$season` function.

## simul$season

This module is used to perform a season where every agent is matched against
other agents a single time and the winners earns token from the looser. Let's
do a single season using the null model where the outcome of a match does not
depend on the amount of token the agents lost or gained during the competitions.


```{r}
res=simul$season(LETTERS[1:8])
res$token
res$M
```

A -1 in the matrix means a loss, a 1 a win, where you must read rowwise. So for instance A was loosing agains B in this season.
The number of tokens is directly related to the number of wins and losses
after a single season. We can use the token to create a new season with non-equal starting conditions.

```{r}
res2=simul$season(LETTERS[1:8],token=res$token)
res2$token
res2$M
```


